---
title: "SurfaceMesh: 3D Surface Meshes Based on CGAL"
author: "Daniel Wollschläger (<wollschlaeger@uni-mainz.de>)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  rmarkdown::html_vignette:
    fig_height: 5
    fig_width: 5
    toc: true
  html_document:
    df_print: paged
    toc: true
  pdf_document:
    fig_height: 5
    fig_width: 5
    toc: true
vignette: >
  %\VignetteIndexEntry{SurfaceMesh}
  %\VignetteDepends{SurfaceMesh}
  %\VignetteKeywords{SurfaceMesh}
  %\VignettePackage{SurfaceMesh}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, echo=FALSE, results="asis"}
source("setup.R")
setupKnitr(autoprint = TRUE)
set.seed(123)

rgl_webshot_org <- Sys.setenv("RGL_USE_WEBSHOT"=FALSE)
```

# Introduction

`SurfaceMesh` is an R package that supports basic processing of 3D surface meshes using as backend the C++ library [`CGAL`](https://www.cgal.org/) for computational geometry via R package [`RcppCGAL`](https://cran.r-project.org/package=RcppCGAL). Features include:

  * Read in and write mesh files in common formats (STL, PLY, OBJ, OFF)
  * Conversion to / from class `mesh3d` from package [`rgl`](https://cran.r-project.org/package=rgl), also compatible with package [`Rvcg`](https://cran.r-project.org/package=Rvcg)
  * Mesh repair
  * Boolean mesh operations on bounded volumes
  * Remeshing
  * Subdivision
  * Surface reconstruction
  * Bounding box
  * Convex hull
  * Distance from points to mesh
  * Centroid (center of mass)
  * Surface area
  * Volume
  * Vertex normals
  * Randomly sampling points on a mesh

The package also calculates surface distance and volume-overlap based similarity metrics for a given pair of 3D surface meshes (see [dkfz metrics reloaded](https://metrics-reloaded.dkfz.de/metric-library)).

# Background

This package includes code adapted from packages [`Boov`](https://github.com/stla/Boov/), [`PolygonSoup`](https://github.com/stla/PolygonSoup/), and [`cgalMeshes`](https://github.com/stla/cgalMeshes/) developed and copyright by [Stéphane Laurent](https://laustep.github.io/stlahblog/). Currently, only a subset of the functionality of these packages is provided in `SurfaceMesh`.

A fork / adaptation of packages [`Boov`](https://github.com/stla/Boov/), [`PolygonSoup`](https://github.com/stla/PolygonSoup/), and [`cgalMeshes`](https://github.com/stla/cgalMeshes/) was carried out as upstream changes to CGAL introduced incompatibilities, and the packages were archived from [CRAN](https://cran.r-project.org/).

The design was chosen such that mesh data resides in R space. This means that for each mesh operation, data is first transferred to the C++ side (using `Rcpp`), converted to a CGAL surface mesh, processed with CGAL functions, and then transferred back to R. The package does not maintain a pointer to a C++ data structure to keep the mesh data there - unlike packages such as [`terra`](https://cran.r-project.org/package=terra) or [`cgalMeshes`](https://github.com/stla/cgalMeshes/). This approach carries a performance penalty, but from an R perspective, it is more straightforward. In particular, there are no serialization issues (saving meshes). Furthermore, memory management is easier.

# Included Data
## Human Hearts

Whole heart delineated by two observers in the same set of CT image slices. Class `CGALmesh`. To display using package `rgl`, first convert to class `mesh3d`.

```{r}
heart1_rgl <- toRGL(dataHeart1)
wire3d(heart1_rgl)
```

```{r}
heart2_rgl <- toRGL(dataHeart2)
wire3d(heart2_rgl)
```

## Geometric forms

Hopf Torus. Class `mesh3d` from package `rgl`.

```{r}
view3d(25, 0, zoom=0.8)
wire3d(dataHopfTorus)
```

Oloid. Class `CGALmesh`. To view using package `rgl`, first convert to class `mesh3d`.

```{r}
oloid_rgl <- toRGL(dataOloid)
view3d(30, 30, zoom=0.8)
wire3d(oloid_rgl)
```

Pentagrammic prism. A list with components `vertices`, `faces`. To view using package `rgl`, first make a `CGALmesh` object and convert to class `mesh3d`.

```{r}
pentaPrism     <- makeMesh(dataPentaPrism, triangulate=TRUE)
pentaPrism_rgl <- toRGL(pentaPrism)
view3d(0, 30, zoom=0.8)
wire3d(pentaPrism_rgl)
```

Septuaginta, i.e. a polyhedron with 72 faces like a sphere based on the drawing by Leonardo da Vinci. Class `mesh3d` from package `rgl`.

```{r}
view3d(30, 30, zoom=0.8)
wire3d(dataSeptuaginta)
```

Truncated icosahedron. A list with components `vertices`, `faces`. To view using package `rgl`, first make a `CGALmesh` object and convert to class `mesh3d`.

```{r}
truncIcosahedron     <- makeMesh(dataTruncIcosahedron, triangulate=TRUE)
truncIcosahedron_rgl <- toRGL(truncIcosahedron)
view3d(30, 30, zoom=0.8)
wire3d(truncIcosahedron_rgl)
```

Tubular knot. Class `mesh3d` from package `rgl`.

```{r}
view3d(0, 0, zoom=0.8)
wire3d(dataTubularKnot)
```

# Create Surface Meshes
## Read Mesh / Polygon Soup Files

```{r}
f_mesh   <- system.file("extdata", "dataHeart3.ply", package="SurfaceMesh")
mesh     <- makeMesh(f_mesh, verbose=FALSE)
mesh_rgl <- toRGL(mesh)

view3d(0, -90, zoom=0.8)
wire3d(mesh_rgl)
```

## Create Surface Mesh Based on `mesh3d` Object

```{r}
dataTubularKnot

mesh <- makeMesh(dataTubularKnot)
mesh
```

## Triangulate

```{r}
mesh     <- makeMesh(dataSeptuaginta, verbose=FALSE)
mesh_rgl <- toRGL(mesh)
isTriangle(mesh)

mesh_tri     <- triangulateMesh(mesh)
mesh_tri_rgl <- toRGL(mesh_tri)

mfrow3d(1, 2)
view3d(0, 30, zoom=0.8)
wire3d(mesh_rgl)
next3d()
view3d(0, 30, zoom=0.8)
wire3d(mesh_tri_rgl)
```

## Vertex normals

```{r}
mesh <- makeMesh(dataSeptuaginta, triangulate=TRUE, normals=TRUE)
mesh
```

Or

```{r}
mesh <- makeMesh(dataSeptuaginta, triangulate=TRUE, normals=FALSE)
mesh

mesh_n <- assignNormals(mesh)
mesh_n
```

# Mesh Repair

See [`CGAL` mesh repair](https://doc.cgal.org/latest/PMP_Mesh_repair/)

## Fill Boundary Holes

Note: Hole is easier to see in interactive version shown with `example(fillBoundaryHoles)`.

```{r}
f_mesh <- system.file("extdata", "dataCornerHole.off", package="SurfaceMesh")
mesh   <- makeMesh(f_mesh, triangulate=TRUE, verbose=TRUE)
mesh_rgl  <- toRGL(mesh)
mesh_fill <- fillBoundaryHoles(mesh, verbose=TRUE)
mesh_fill_rgl <- toRGL(mesh_fill)

mfrow3d(1, 2)
view3d(30, -45, zoom=0.8)
wire3d(mesh_rgl)
next3d()
view3d(30, -45, zoom=0.8)
wire3d(mesh_fill_rgl)
```

## Remove Self-Intersections

TODO: Use a mesh with self-intersections.

```{r}
mesh     <- makeMesh(dataPentaPrism, triangulate=TRUE)
mesh_nsi <- removeSelfIntersections(mesh)
getVolume(mesh_nsi)
```

## Randomly Sample Points on a Mesh

```{r}
mesh      <- makeMesh(dataPentaPrism, triangulate=TRUE)
ptsOnMesh <- samplePoints(mesh, ptsOnFaces=2L)
head(ptsOnMesh)
```

# Boolean Mesh Operations on Bounded Volumes

See [`CGAL` Boolean operations](https://doc.cgal.org/latest/PMP_Boolean_operations/)

## Union

```{r}
# mesh one: a cube
mesh1_rgl  <- cube3d()   # (from the rgl package)
mesh2_rgl  <- translate3d(cube3d(), 1, 1, 1)
mesh_u     <- boolUnion(list(mesh1_rgl, mesh2_rgl))
mesh_u_rgl <- toRGL(mesh_u)

shade3d(mesh_u_rgl, color="red")
```

## Difference

```{r}
mesh1_rgl  <- cube3d()   # (from the rgl package)
mesh2_rgl  <- translate3d(cube3d(), 1, 1, 0)
mesh_d     <- boolDifference(mesh1_rgl, mesh2_rgl)
mesh_d_rgl <- toRGL(mesh_d)

shade3d(mesh1_rgl,  color="yellow", alpha=0.2)
shade3d(mesh2_rgl,  color="cyan",   alpha=0.2)
shade3d(mesh_d_rgl, color="red")
```

## Intersection

```{r}
mesh1      <- makeMesh(dataTruncIcosahedron, triangulate=TRUE)
mesh1_rgl  <- toRGL(mesh1)
mesh2_rgl  <- translate3d(cube3d(), 2, 0, 0)
mesh2      <- makeMesh(mesh2_rgl, triangulate=TRUE)
mesh_i     <- boolIntersection(list(mesh1, mesh2))
mesh_i_rgl <- toRGL(mesh_i)

shade3d(mesh1_rgl,  color="yellow", alpha=0.2)
shade3d(mesh2_rgl,  color="cyan",   alpha=0.2)
shade3d(mesh_i_rgl, color="red")
```

# Remeshing

See [`CGAL` remeshing](https://doc.cgal.org/latest/PMP_Remeshing/)

## Isotropic

```{r}
mesh         <- makeMesh(dataTruncIcosahedron, triangulate=TRUE)
mesh_rgl     <- toRGL(mesh)
mesh_rem     <- remeshIsotropic(mesh, targetEdgeLen=1)
mesh_rem_rgl <- toRGL(mesh_rem)

mfrow3d(1, 2)
view3d(0, 0, zoom=0.7)
wire3d(mesh_rgl)
next3d()
view3d(0, 0, zoom=0.7)
wire3d(mesh_rem_rgl)
```

## Triangulated Surface Mesh Simplification

```{r}
mesh       <- makeMesh(dataHopfTorus, triangulate=TRUE)
mesh_rgl   <- toRGL(mesh)
mesh_s     <- remeshSimplify(mesh, method="LT-R", stopUeRatio=0.1)
mesh_s_rgl <- toRGL(mesh_s)

mfrow3d(1, 2)
view3d(0, 0, zoom=0.7)
wire3d(mesh_rgl)
next3d()
view3d(0, 0, zoom=0.7)
wire3d(mesh_s_rgl)
```

## Smoothing

```{r}
mesh       <- dataHeart1
mesh_rgl   <- toRGL(mesh)
mesh_s     <- remeshSmooth(mesh, nIter=5, time=1)
mesh_s_rgl <- toRGL(mesh_s)

mfrow3d(1, 2)
view3d(0, -90, zoom=0.7)
wire3d(mesh_rgl)
next3d()
view3d(0, -90, zoom=0.7)
wire3d(mesh_s_rgl)
```

# Subdivision

See [`CGAL subdivision`](https://doc.cgal.org/latest/Subdivision_method_3/)

## Catmull-Clark

```{r}
mesh        <- makeMesh(dataPentaPrism, triangulate=TRUE)
mesh_rgl    <- toRGL(mesh)
mesh_cc     <- subdivideCatmullClark(mesh, nIter=2)
mesh_cc_rgl <- toRGL(mesh_cc)

mfrow3d(1, 2)
view3d(0, 30, zoom=0.8)
wire3d(mesh_rgl)
next3d()
view3d(0, 30, zoom=0.8)
wire3d(mesh_cc_rgl)
```

## Doo-Sabin

```{r}
mesh        <- makeMesh(dataPentaPrism, triangulate=TRUE)
mesh_rgl    <- toRGL(mesh)
mesh_ds     <- subdivideDooSabin(mesh, nIter=2)
mesh_ds_rgl <- toRGL(mesh_ds)

mfrow3d(1, 2)
view3d(0, 30, zoom=0.8)
wire3d(mesh_rgl)
next3d()
view3d(0, 30, zoom=0.8)
wire3d(mesh_ds_rgl)
```

## Sqrt3

```{r}
mesh        <- makeMesh(dataPentaPrism, triangulate=TRUE)
mesh_rgl    <- toRGL(mesh)
mesh_s3     <- subdivideSqrt3(mesh, nIter=2)
mesh_s3_rgl <- toRGL(mesh_s3)

mfrow3d(1, 2)
view3d(0, 30, zoom=0.8)
wire3d(mesh_rgl)
next3d()
view3d(0, 30, zoom=0.8)
wire3d(mesh_s3_rgl)
```

## Loop

```{r}
mesh       <- makeMesh(dataPentaPrism, triangulate=TRUE)
mesh_rgl   <- toRGL(mesh)
mesh_l     <- subdivideLoop(mesh, nIter=2)
mesh_l_rgl <- toRGL(mesh_l)

mfrow3d(1, 2)
view3d(0, 30, zoom=0.8)
wire3d(mesh_rgl)
next3d()
view3d(0, 30, zoom=0.8)
wire3d(mesh_l_rgl)
```

# Surface Reconstruction
## AFS

See [`CGAL` AFS](https://doc.cgal.org/latest/Advancing_front_surface_reconstruction/)

```{r}
mesh         <- dataHeart1
mesh_rgl     <- toRGL(mesh)
mesh_afs     <- reconstructAFS(mesh[["vertices"]], jetSmoothing=30)
mesh_afs_rgl <- toRGL(mesh_afs)

mfrow3d(1, 2)
view3d(0, -90, zoom=0.7)
wire3d(mesh_rgl)
next3d()
view3d(0, -90, zoom=0.7)
wire3d(mesh_afs_rgl)
```

## SSS

See [`CGAL` SSS](https://doc.cgal.org/latest/Scale_space_reconstruction_3/)

```{r}
mesh     <- dataHeart1
mesh_rgl <- toRGL(mesh)
mesh_sss <- reconstructSSS(mesh[["vertices"]],
                           scaleIterations=1L,
                           forceManifold  =TRUE,
                           neighbors      =6L)
mesh_sss_rgl <- toRGL(mesh_sss)

mfrow3d(1, 2)
view3d(0, -90, zoom=0.7)
wire3d(mesh_rgl)
next3d()
view3d(0, -90, zoom=0.7)
wire3d(mesh_sss_rgl)
```

## Poisson

See [`CGAL` SSS](https://doc.cgal.org/latest/Poisson_surface_reconstruction_3/)

```{r}
mesh     <- makeMesh(dataHopfTorus)
mesh_rgl <- toRGL(mesh)
mesh_psr <- reconstructPoisson(mesh[["vertices"]],
                               normalsFun=getNormalsFun(6L),
                               smAngle=10,
                               smRadius=3,
                               smDistance=0.3)
mesh_psr_rgl <- toRGL(mesh_psr)

mfrow3d(1, 2)
wire3d(mesh_rgl)
next3d()
wire3d(mesh_psr_rgl)
```

## Alpha Wrapping

See [`CGAL` Alpha Wrapping](https://doc.cgal.org/latest/Alpha_wrap_3/)

```{r}
mesh        <- dataHeart1
mesh_rgl    <- toRGL(dataHeart1)
mesh_aw     <- alphaWrap(mesh,
                         alphaRel =5,
                         offsetRel=300)
mesh_aw_rgl <- toRGL(mesh_aw)

mfrow3d(1, 2)
view3d(0, -90, zoom=0.7)
wire3d(mesh_rgl)
next3d()
view3d(0, -90, zoom=0.7)
wire3d(mesh_aw_rgl)
```

# Bounding Shapes
## Axis-parallel Bounding Box

```{r}
mesh     <- dataOloid
mesh_rgl <- toRGL(mesh)
bb       <- getBoundingBox(mesh)
bb_rgl   <- toRGL(bb)

view3d(0, 30, zoom=0.9)
wire3d(mesh_rgl)
wire3d(bb_rgl)
```

## Optimal (Oriented) Bounding Box

```{r}
mesh     <- dataHeart1
mesh_rgl <- toRGL(mesh)
obb      <- getBoundingBoxOptimal(mesh)
obb_rgl  <- toRGL(obb[["mesh"]])

view3d(0, -90, zoom=0.8)
wire3d(mesh_rgl)
wire3d(obb_rgl)
```

## Convex Hull

```{r}
mesh     <- makeMesh(dataHopfTorus)
mesh_rgl <- toRGL(mesh)
hull     <- getConvexHull(mesh[["vertices"]])
hull_rgl <- toRGL(hull)

mfrow3d(1, 2)
view3d(0, 0, zoom=0.8)
wire3d(mesh_rgl)
next3d()
view3d(0, 0, zoom=0.8)
wire3d(hull_rgl)
```

# Measurements
## Centroid (Center of Mass)

```{r}
mesh <- makeMesh(dataPentaPrism, triangulate=TRUE)
getCentroid(mesh)
```

## Surface Area

```{r}
mesh <- makeMesh(dataTruncIcosahedron, triangulate=TRUE)
getArea(mesh)
```

## Volume

```{r}
mesh <- makeMesh(dataSeptuaginta, triangulate=TRUE)
getVolume(mesh)
```

# Distace and Agreement Measures
## Euclidean Distance from Points to a Mesh

```{r}
mesh   <- makeMesh(dataPentaPrism, triangulate=TRUE)
points <- matrix(2*runif(3*4), ncol=3)
getDistance(mesh, points)
```

## Surface Distance between two Meshes
### Distance between the two respective centers of mass (DCOM)

```{r}
getDCOM(dataHeart1, dataHeart2)
```

### Hausdorff distance

See [`CGAL` Hausdorff distance](https://doc.cgal.org/latest/Polygon_mesh_processing/index.html#PMPDistance)

Approximate

```{r}
getHausdorff(dataHeart1, dataHeart2, symmetric=TRUE)
```

Bounded error

```{r}
getHausdorff(dataHeart1, dataHeart2, symmetric=TRUE, errorBound=0.001)
```

Quantile (e.g., 'HD95')

```{r}
getHausdorffQuantile(dataHeart1, dataHeart2, p=0.95)
```

### Average symmetric surface distance (ASSD) and Root Mean Squared Error (RMSE)

See [dkfz metric reloaded: ASSD](https://metrics-reloaded.dkfz.de/metric-library/assd)

```{r}
sd_heart <- getSurfaceDist(dataHeart1, dataHeart2, p=0.95, ptsOnFaces=100L)
sd_heart[["ASSD"]]
sd_heart[["RMSE"]]
```

## Volume-Overlap Based Similarity Measures

  * [dkfz metric reloaded: Jaccard Similarity Coefficient](https://metrics-reloaded.dkfz.de/metric-library/intersection_over_union) (JSC, aka 'Intersection over Union', IoU)
  * [dkfz metric reloaded: Dice Similarity Coefficient](https://metrics-reloaded.dkfz.de/metric-library/dsc) (DSC)

```{r}
mesh1_rgl <- cube3d()      # (from the rgl package)
mesh2_rgl <- translate3d(cube3d(), 1, 1, 1)
jsc_dsc   <- getJSCDSC(mesh1_rgl, mesh2_rgl)
jsc_dsc[["JSC"]]
jsc_dsc[["DSC"]]
```

```{r echo=FALSE, results="asis"}
Sys.setenv("RGL_USE_WEBSHOT"=rgl_webshot_org)
```
