---
title: "SurfaceMesh: Example Meshes"
author: "Daniel Wollschläger (<wollschlaeger@uni-mainz.de>)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  rmarkdown::html_vignette:
    fig_height: 5
    fig_width: 5
    toc: true
  html_document:
    df_print: paged
    toc: true
  pdf_document:
    fig_height: 5
    fig_width: 5
    toc: true
vignette: >
  %\VignetteIndexEntry{SurfaceMesh Examples}
  %\VignetteDepends{SurfaceMesh}
  %\VignetteKeywords{SurfaceMesh}
  %\VignettePackage{SurfaceMesh}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, echo=FALSE, results="asis"}
source("setup.R")
setupKnitr(autoprint=TRUE)
set.seed(123L)

rgl_webshot_org <- Sys.setenv("RGL_USE_WEBSHOT"=FALSE)
```

# Geometric Meshes

The geometric meshes included in package `SurfaceMesh` were originally part of R packages [`SurfaceReconstruction`](https://github.com/stla/SurfaceReconstruction/) and [`cgalMeshes`](https://github.com/stla/cgalMeshes/), developed and copyright by [Stéphane Laurent](https://laustep.github.io/stlahblog/). Please see the following directory for the `.R` scripts for how to generate the meshes.

```{r}
## package installation directory depends on system
d <- system.file("data-raw", package="SurfaceMesh")
list.files(d)
```

## Dupin Cyclide

```{r}
view3d(0, -35, zoom=0.8)
shade3d(dataCyclide, color="orangered")
```

## Entzensberger Star

```{r}
view3d(-55, -20, zoom=0.8)
shade3d(dataEntzenStar, col="steelblue2")
```

## Hopf Torus

```{r}
view3d(35, -20, zoom=0.8)
shade3d(toRGL(dataHopfTorus), color="darkred")
```

## ICN5D Eight

```{r}
view3d(35, -20, zoom=0.8)
shade3d(dataICN5Deight, color="violetred")
```

## Oloid

```{r}
view3d(20, 15, zoom=0.8)
shade3d(toRGL(dataOloid), color="darkviolet")
```

## OrthoCircle

[Paul Bourke OrthoCircle](https://paulbourke.net/geometry/orthocircle/)

```{r}
view3d(-15, -10, zoom=0.8)
shade3d(dataOrthoCircle, color="darkolivegreen4")
```

## Pentagrammic Prism

```{r}
meshPentaPrism <- makeMesh(dataPentaPrism, triangulate=TRUE)
view3d(-15, -15, zoom=0.8)
shade3d(toRGL(meshPentaPrism), color="forestgreen")
```

## Septuaginta

```{r}
view3d(-10, -20, zoom=0.8)
shade3d(dataSeptuaginta, color="orangered")
```

## Solid Möbius strip

```{r}
view3d(-10, -20, zoom=0.8)
shade3d(dataSolidMobiusStrip, color="darkred")
```

## Sphere

```{r}
view3d(-25, -45, zoom=0.8)
shade3d(dataSphere, color="darkgoldenrod")
wire3d(dataSphere)
```

## Spider Cage

```{r}
view3d(-25, -45, zoom=0.8)
shade3d(dataSpiderCage, color="mediumseagreen")
```

## Toroidal Helix

```{r}
view3d(15, -20, zoom=0.8)
shade3d(dataToroHelix, color="violetred")
```

## Torus

```{r}
view3d(-20, -20, zoom=0.75)
shade3d(dataTorus, color="green")
wire3d(dataTorus)
```

## Truncated Icosahedron

```{r}
meshTruncIcos <- makeMesh(dataTruncIcosahedron, triangulate=TRUE)
view3d(-15, -10, zoom=0.8)
shade3d(toRGL(meshTruncIcos), color="wheat2")
```

## Tubular Knot

```{r}
view3d(-15, -15, zoom=0.8)
shade3d(dataTubularKnot, color="royalblue3")
```

## Corner

```{r}
f_mesh   <- system.file("extdata", "dataCorner.off", package="SurfaceMesh")
mesh     <- makeMesh(f_mesh, verbose=FALSE)
mesh_rgl <- toRGL(mesh)

view3d(85, -65, zoom=0.8)
shade3d(mesh_rgl, col="turquoise4")
```

# Human Heart Contoured from CT Images
## Heart by Observer 1

```{r}
view3d(-5, -90, zoom=0.8)
shade3d(toRGL(dataHeart1), color="tomato")
```

## Heart by Observer 2

```{r}
view3d(-5, -90, zoom=0.8)
shade3d(toRGL(dataHeart2), color="gray")
```
